---
title: "Museums and Nature Centers"
output: html_notebook
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

There are thousands of museums, aquariums, and zoos across the United States. In this project, we’ll take a look at the distribution of these institutions by geographic region, type, and revenue.

Our data is compiled from administrative records from the Institute of Museum and Library Services, IRS records, and private foundation grantmaking records. This data reflects the status of each institution as of 2013. For each institution, we have information on its name, type, and location. Each institution also has a parent organization – for example, if a museum housed at a university, its parent organization is the university where it resides. Financial data on annual revenue is available at the parent organization level. We’ll be creating several different data visualizations in this project. 


```{r data, message=FALSE}
# install packages
install.packages(c('plotrix'))

# load libraries
library(dplyr)
library(ggplot2)
library(stringr)
library(tidyr)
library(plotrix)

```

## Data Exploration

```{r Import .txt file and write to csv}
#Import txt file in R
museums = read.table("museums.txt", sep=",", header=TRUE)

# storing this dataframe in a csv file
write.csv(museums, file = 'museums.csv', row.names = FALSE)
```



```{r load csv, message=FALSE}
# Load file as data frame
museums_df <- read.csv("museums.csv")
```


```{r inspect, message=FALSE}
# Inspect data frame
head(museums_df)
```

## Museums by Type

```{r barplot, message=FALSE}
# Create and print bar plot by type
museum_type <- ggplot(museums_df, aes(x=Museum.Type)) +
  geom_bar() + 
  # customize x axis, set wrap_format() to 8
  scale_x_discrete(labels = scales::wrap_format(8)) 
museum_type
```

```{r barplot_museum, message=FALSE}
# Create and print bar plot by museum vs non-museum
museum_class <- ggplot(museums_df, aes(x=Is.Museum)) +
  geom_bar() +
  # rename x axis
  scale_x_discrete(labels = c("FALSE" = "Non-Museum", "TRUE" = "Museum"))
museum_class
```

```{r barplot_type, message=FALSE}
# Filter data frame to select states
museums_states <- museums_df %>%
  filter(
    grepl('IL|CA|NY', 
    State..Administrative.Location.))
museums_states
# Create and print bar plot with facets
museum_facet <- ggplot(museums_states, aes(x=Is.Museum)) +
  geom_bar() +
  facet_grid(
    cols = vars(State..Administrative.Location.)) +
  # rename x axis
  scale_x_discrete(labels = c("FALSE" = "Non-Museum", "TRUE" = "Museum"))
museum_facet
```

```{r barplot_stack, message=FALSE}
# Create and print stacked bar plot
museum_stacked <- 
  ggplot(museums_df, 
    aes(
      x = factor(Region.Code..AAM.), 
      fill = Is.Museum)) + 
  geom_bar(position = "fill") + 
  scale_x_discrete(
    labels = c(
      "1"="New England",
      "2"="Mid-Atlantic", 
      "3"="Southeastern",
      "4"="Midwest", 
      "5"="Mountain Plains", 
      "6"="Western")) + 
  scale_y_continuous(
    labels = scales::percent_format()) +
  scale_fill_discrete(
    labels = c(
      "TRUE" = "Museum", 
      "FALSE" = "Non-Museum")) + 
  labs(
    title = "Museum Types by Region",
    x = "Region",
    y = "Percentage of Total",
    fill = "Type"
  )
  
museum_stacked
```

## Museums by Revenue

```{r process, message=FALSE}
# Filter data frame
museums_revenue_df <- museums_df %>%
  distinct(Legal.Name, .keep_all = TRUE) %>%
  filter(Annual.Revenue > 0)

# Filter for only small museums
museums_small_df <- museums_revenue_df %>%
  filter(Annual.Revenue < 1e6)

# Filter for only large museums
museums_large_df <- museums_revenue_df %>%
  filter(Annual.Revenue > 1e9)
```

```{r histogram, message=FALSE}
# Create and print histogram
revenue_histogram <- 
  ggplot(museums_small_df, 
    aes(x = Annual.Revenue)) + 
  geom_histogram(binwidth = 1e4) +
  scale_x_continuous(
    labels = scales::dollar_format())

revenue_histogram
```

```{r boxplot, message=FALSE}
# Create and print boxplot
revenue_boxplot <- 
  ggplot(museums_large_df, 
    aes(
      x = factor(Region.Code..AAM.), 
      y = Annual.Revenue)) + 
  geom_boxplot() + 
  coord_cartesian(ylim = c(1e9, 3e10)) +
  scale_y_continuous(
    labels = function(x) paste0("$", x/1e9, "B")) +
  scale_x_discrete(
    labels = c(
      "1"="New England",
      "2"="Mid-Atlantic", 
      "3"="Southeastern",
      "4"="Midwest", 
      "5"="Mountain Plains", 
      "6"="Western"))

revenue_boxplot
```

```{r mean, message=FALSE}
# Create and print bar plot with means
revenue_barplot <- 
  ggplot(museums_revenue_df, 
    aes(
      x = factor(Region.Code..AAM.), 
      y = Annual.Revenue)) + 
  geom_bar(stat = "summary", fun = "mean") + 
  scale_y_continuous(
    labels = scales::dollar_format()) + 
  scale_x_discrete(
    labels = c(
      "1"="New England",
      "2"="Mid-Atlantic", 
      "3"="Southeastern",
      "4"="Midwest", 
      "5"="Mountain Plains", 
      "6"="Western")) +
  labs(
    title = "Mean Annual Revenue by Region", 
    y = "Mean Annual Revenue", 
    x = "Region")
  
revenue_barplot
```

```{r mean_errorbar, message=FALSE}
# Calculate means and standard errors
museums_error_df <- museums_revenue_df %>%
  group_by(Region.Code..AAM.) %>%
  summarize(
    Mean.Revenue = mean(Annual.Revenue), 
    Mean.SE = std.error(Annual.Revenue)) %>%
  mutate(
    SE.Min = Mean.Revenue - Mean.SE, 
    SE.Max = Mean.Revenue + Mean.SE)

# Create and print bar plot with means and standard errors
revenue_errorbar <- 
  ggplot(museums_error_df, 
    aes(
      x = factor(Region.Code..AAM.), 
      y = Mean.Revenue)) + 
  geom_bar(stat = "identity") +
  geom_errorbar(
    aes(ymin = SE.Min, ymax=SE.Max), width=0.2) +
  scale_y_continuous(
    labels = scales::dollar_format()) + 
  scale_x_discrete(
    labels = c(
      "1"="New England",
      "2"="Mid-Atlantic", 
      "3"="Southeastern",
      "4"="Midwest", 
      "5"="Mountain Plains", 
      "6"="Western")) +
  labs(
    title = "Mean Annual Revenue by Region", 
    y = "Mean Annual Revenue", 
    x = "Region")
  
revenue_errorbar
```
